---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scimo

```{r, echo = FALSE}
version <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
version <- gsub('-', '.', version)
```

<!-- badges: start -->
[![experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![packageversion](https://img.shields.io/badge/version-`r version`-orange.svg)](commits/master)
[![R-CMD-check](https://github.com/abichat/scimo/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/abichat/scimo/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**scimo** provides extra recipes steps for dealing with omics data, but not only.


## Installation

You can install **scimo** from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("abichat/scimo")
```

## Example

```{r data, message=FALSE}
library(scimo)

data("pedcan_expression")
pedcan_expression
```

```{r seed, echo=FALSE}
set.seed(42)
```

```{r example}
rec <-
  recipe(pedcan_expression) %>% 
  update_role(disease, new_role = "outcome") %>% 
  update_role(-disease, new_role = "predictor") %>% 
  update_role(cell_line, new_role = "ID") %>% 
  step_select_cv(all_numeric_predictors(), n_kept = 3000) %>% 
  step_select_kruskal(all_numeric_predictors(), cutoff = 0.05,
                      outcome = "disease", correction = "fdr") %>% 
  prep()

rec

juice(rec)

tidy(rec, 1)
tidy(rec, 2)
```


## Notes

### `protection stack overflow` error

If you have a very large dataset, you may encounter this error: 

```{r error, error=TRUE}
recipe(disease ~ ., data = pedcan_expression) %>% 
    step_select_cv(all_numeric_predictors(), prop_kept = 0.1) 
```

It is linked to [how **R** handles many variables in formulas](https://github.com/tidymodels/recipes/issues/467). To solve it, pass only the dataset to `recipe()` and manually update roles with `update_role()`, like in the example above.


### Steps for variable selection

Like [**colino**](https://github.com/stevenpawley/colino), **scimo** proposes 3 arguments for variable selection steps based on a statistic: `n_kept`, `prop_kept` and `cutoff`.

* `n_kept` and `prop_kept` deal with how many variables will be kept in the preprocessed dataset, based on an exact count of variables or a proportion relative to the original dataset. They are mutually exclusive.

* `cutoff` removes variables whose statistic is below (or above, depending on the step) it. It could be used alone or in addition to the two others. 



